\documentclass[12pt]{article}
%\usepackage{html}
\usepackage{hyperref}
\title{CS 571\\Quiz 2}
%\addtolength{\topmargin}{-2cm}
%\addtolength{\topskip}{-2cm}
%\addtolength{\oddsidemargin}{-2cm}
%\addtolength{\evensidemargin}{-2cm}
%\addtolength{\textheight}{2cm}
%\addtolength{\textwidth}{2cm}
%\addtolength{\footskip}{-1cm}
\date{}
\begin{document}
\maketitle

\begin{flushleft}
\textbf{Oct 5}\hfill\textbf{Closed book}\\
\textbf{15 points}\hfill\textbf{Closed notes}\\

\vspace{0.5cm}

\textbf{Important Reminder}: As per the course Academic Honesty
Statement, cheating of any kind will minimally result in receiving an
F letter grade for the entire course.


\end{flushleft}

For each of the following questions, select a \textbf{single}
alternative on the grid-sheet.  Please ensure that you have filled-in
your name and B-number in the bubbles on the provided grid-sheet.

There are 7 questions with 2-points per question; there is 1-point
for submitting the quiz.

\begin{enumerate}

\item Which of the following languages over the vocabulary of
  parentheses $\{$\verb@(@, \verb@)@$\}$ is not expressible using
  standard regular expressions?

\begin{enumerate}

\item Strings whose length is exactly 3.  Examples include \verb@()(@,
  \verb@)()@ and \verb@())@.
  
\item Strings of even length.  Examples include
  the empty string, \verb@((@, \verb@()()@ and \verb@(())@.

\item Strings of length less-than-or-equal-to 4 which consist of
  balanced parentheses. 
    Examples include the empty string, \verb@()@ and \verb@()()@.

\item Strings of even length containing 1-or-more \verb@(@'s
  followed by 0-or-more \verb@)@'s.  Examples include \verb@((@,
  \verb@()))@ and \verb@(())@.

\item Strings of even length which consist of balanced parentheses.
  Examples include the empty string, \verb@()@ and \verb@()(())@.

\end{enumerate}

\item Given the following CFG over the set of terminals $\{$\verb@NUM@, \verb@ID@, \verb@','@, \verb@';'@$\}$:

\begin{verbatim}
list
  : ID tail
  ;
tail
  : ',' NUM tail
  | ',' ID tail
  | /* empty */
  | ';'
  ;
\end{verbatim}

Which one of the following describes the language defined by the above
CFG most precisely?

\begin{enumerate}

\item Lists of \verb@ID@'s and \verb@NUM@'s 
  separated by \verb@','@ and terminated by \verb@';'@.
  
\item Lists of \verb@ID@'s and \verb@NUM@'s starting with a \verb@ID@,
  separated by \verb@','@ and optionally terminated by \verb@';'@.

\item Lists of \verb@ID@'s and \verb@NUM@'s 
  separated by \verb@','@.
  
\item Lists of \verb@ID@'s and \verb@NUM@'s 
  separated by \verb@','@ and optionally terminated by \verb@';'@.
  
\item Lists of \verb@ID@'s and \verb@NUM@'s starting with a \verb@ID@,
  separated by \verb@','@ and terminated by \verb@';'@.
    
\end{enumerate}

\item In Javascript, \textit{hoisting} refers to:
\begin{enumerate}

\item Moving \verb@var@ declarations to the start
  of a block.

\item Moving \verb@var@ declarations to the start
  of a function.

\item Moving \verb@var@ declarations to the \verb@window@ object.

\item Moving \verb@let@ declarations to the start
  of a block.

\item Moving \verb@let@ declarations to the start
  of a function.

\end{enumerate}

\newpage


\item Heap allocation is the only alternative for

\begin{enumerate}

\item Entities having a lifetime equal to that of the entire
  program.

\item Entities having a lifetime equal to that of a function activation.

\item Entities having a lifetime equal to that of a block activation.

\item Entities which are function parameters.

\item Entities which have indeterminate lifetime.

\end{enumerate}


\item Temporary variables within a stack frame refer to

\begin{enumerate}

  \item Variables declared to be temporary by the programmer.

  \item Variables introduced by the programmer to implement the
    exchange of the values of two variables.

  \item Variables which contain references to heap data.

  \item Variables introduced by the compiler to hold
    the variables of a function while it is active.

  \item Variables introduced by the compiler to hold intermediate
    values computed while evaluating an expression.
         
    
\end{enumerate}

\newpage

\item Assuming that all declarations introduced using \verb@var@ in
  the following pseudo-code are dynamically scoped, what will be
  the output of the following program?

\begin{verbatim}
var a = 11;

f() { var a = 22; h(); }

h() { print a; }

print a; f(); print a;
\end{verbatim}

\begin{enumerate}

\item \verb@11 11 11@

\item \verb@11 22 11@

\item \verb@11 22 22@

\item \verb@11 11 22@

\item \verb@22 11 22@
    
\end{enumerate}

\newpage

\item Which one of the following statements is false?

\begin{enumerate}

\item Scheme is dynamically scoped.

\item In some languages, there can be \textit{holes} within the scope
  of a variable.

\item A context-free grammar can describe nested constructs.

\item In C, the declaration \verb@struct S;@ can be used to make a
  forward-reference to a structure type.

\item Static allocation cannot be the sole allocation strategy in
  the presence of recursive functions.

\end{enumerate}

\end{enumerate}

\end{document}
